---
title: "My First Project (Kinda)"
author: "Joshua Matamoros"
date: "2026-8-20"
categories: [project, code]
---

Hello! Welcome to my first project (kinda). This blog will demonstrate some skills that I have learned after taking a week long course for R.

Down below I have coded some graphs involving data about vehicles and their mileage depending on whether its being driven in city conditions or highway conditions.

```{r}
library(ggplot2)

mpg

ggplot(
  data = mpg,
  mapping = aes(
    x = cty,
    y = hwy,
    color = year,
  ) 
) + geom_point(position = "jitter", alpha = 0.6) 
```

I have also included a data table which mirrors the information displayed on the graph.

This graph involves minimal coding, only really focusing on how the mileage performs based on the year of the vehicle.

There are several ways that this graph can be improved.

```{r}

ggplot(
  data = mpg,
  mapping = aes(
    x = cty,
    y = hwy,
    color = year,
  ) 
) + geom_point(position = "jitter", alpha = 0.4)  +
  labs(
    title = "MPG while in the city vs highway",
    x = "city fuel economy (mpg)",
    y = "Highway fuel economy (mpg)",
    ) +  scale_y_continuous(breaks = seq(0, 40, by = 5)) +
    scale_x_continuous(breaks = seq(0, 40, by = 5)) 
```

Some ways that this graph was improved were the additions of labels on the x-axis and the y-axis. A title was also added, and more increments were added to understand the graph more clearly.

For the last graph, I will be implementing most of the skills I have learned in this course for graphing.

```{r}

ggplot(
  data = mpg,
  mapping = aes(
    x = cty,
    y = hwy,
    color = year,
  ) 
) + geom_point(position = "jitter", alpha = 0.4)  +
  labs(
    title = "MPG while in the City vs Highway",
    subtitle = "Additional comparison of years of vehicles",
    x = "City Fuel Economy (mpg)",
    y = "Highway fuel Economy (mpg)",
  ) +  scale_y_continuous(breaks = seq(0, 40, by = 5)) +
  scale_x_continuous(breaks = seq(0, 40, by = 5)) +
  facet_wrap(~cyl)
```
